The Mass Distribution of Stellar-Mass Black Holes

### 2010-06-30 paper.tex
Typos in prior for Gaussian, Power-Law priors and Exponential likelihood.
Added Ilya to author list, grants in ack.

\author{Will M. Farr \and Niharika Sravan} \affil{Northwestern
  University Center for Interdisciplinary Exploration and Research in
  Astrophysics\\2145 Sheridan Rd., Evanston, IL 60208}
\email{w-farr@northwestern.edu, niharika.sravan@gmail.com}

@@

\author{Vicky Kalogera} \affil{Northwestern University Center for
  Interdisciplinary Exploration and Research in Astrophysics\\2145
  Sheridan Rd., Evanston, IL 60208} \email{vicky@northwestern.edu}

@@

Many astrophysical distributions are power-laws.  Let us assume that
the BH mass distribution is given by
\begin{equation}
  \label{eq:power-law-dist}
  p(m|\theta) = p(m|\{\Mmin, \Mmax, \alpha \}) =
  \begin{cases}
    A m^\alpha & \Mmin \leq m \leq \Mmax \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The normalizing constant $A$ is given by 
\begin{equation}
  A = \frac{1+\alpha}{\Mmax^{1+\alpha} - \Mmin^{1+\alpha}}.
\end{equation}
To perform the Bayesian analysis, we need to impose a prior on the
parameters $\Mmin$, $\Mmax$, $\alpha$.  We choose to impose a broad
prior that represents our significant uncertainty in the appropriate
values of these parameters.  We use
\begin{equation}
  p(\theta) = p(\{\Mmin, \Mmax, \alpha\}) = 
  \begin{cases}
    \frac{1}{2} \frac{1}{40^2}\frac{1}{20} & 0 \leq \Mmin \leq \Mmax
    \leq 40, \quad -12 \leq \alpha \leq 8 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where the masses are measured in solar masses.  This prior is uniform
in the masses between 0 and $40 \Msun$, and uniform in power-law slope
between $-12$ and $8$.

@@

Theoretical arguments in \citet{Fryer2001} suggest that the black-hole
mass distribution may be well-represented by a decaying exponential
with a minimum mass:
\begin{equation}
  \label{eq:exp-def}
  p(m|\theta) = p(m|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{m}{M_0}
    \right] & M_0 \leq m \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose again a broad, flat prior
\begin{equation}
  p(\{\Mmin, M_0 \}) = 
  \begin{cases}
    \frac{1}{40^2} & 0 \leq \Mmin, M_0 \leq 40 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
with $\Mmin$ and $M_0$ measured in solar masses.

@@

To look for a second peak in the black-hole mass distribution, we also
examined a two-Gaussian model:
\begin{multline}
  p(m|\theta) = p(m|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right],
\end{multline}
where we restrict $\mu_1 < \mu_2$ with our broad, flat prior:
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    \frac{1}{2} \frac{1}{40^4} & 0 \leq \mu_1 \leq \mu_2 \leq 40,
    \quad 0 \leq \sigma_1, \sigma_2 \leq 40, \quad 0 \leq \alpha \leq
    1 \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The one-Gaussian model is a subset of the Gaussian model: when
$\alpha$ is small, then the presence of the first Gaussian is
irrelevant, and the second gives the same mass distribution as a
single Gaussian; when $\alpha$ is large, the reverse situation again
reproduces the single-Gaussian model; and when $\alpha$ is in the
middle of its range, the two means and standard deviations can be
nearly equal, reproducing again the single-Gaussian model.  This is
precisely what we observe from our MCMC samples in the two-Gaussian
model.  In Section \ref{sec:model-selection}, we will quantify more
precisely how much redundancy we have in this model relative to the
one-Gaussian model, but we should emphasize here that there is no
evidence of a significant contribution from a second peak in the
two-Gaussian model.

@@

This work was supported by grant XXX.  WMF would like to thank Ilya
Mandel for illuminating discussions and encouragement.

>>>>

\author{Will M. Farr \and Niharika Sravan \and Ilya Mandel}

\affil{Northwestern University Center for Interdisciplinary
  Exploration and Research in Astrophysics\\2145 Sheridan Rd.,
  Evanston, IL 60208}

\email{w-farr@northwestern.edu, niharika.sravan@gmail.com}

@@

\author{Vicky Kalogera}

\affil{Northwestern University Center for Interdisciplinary
  Exploration and Research in Astrophysics\\2145 Sheridan Rd.,
  Evanston, IL 60208}

 \email{vicky@northwestern.edu}

@@

Many astrophysical distributions are power-laws.  Let us assume that
the BH mass distribution is given by
\begin{equation}
  \label{eq:power-law-dist}
  p(m|\theta) = p(m|\{\Mmin, \Mmax, \alpha \}) =
  \begin{cases}
    A m^\alpha & \Mmin \leq m \leq \Mmax \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The normalizing constant $A$ is given by 
\begin{equation}
  A = \frac{1+\alpha}{\Mmax^{1+\alpha} - \Mmin^{1+\alpha}}.
\end{equation}
To perform the Bayesian analysis, we need to impose a prior on the
parameters $\Mmin$, $\Mmax$, $\alpha$.  We choose to impose a broad
prior that represents our significant uncertainty in the appropriate
values of these parameters.  We use
\begin{equation}
  p(\theta) = p(\{\Mmin, \Mmax, \alpha\}) = 
  \begin{cases}
    2 \frac{1}{40^2}\frac{1}{20} & 0 \leq \Mmin \leq \Mmax
    \leq 40, \quad -12 \leq \alpha \leq 8 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where the masses are measured in solar masses.  This prior is uniform
in the masses between 0 and $40 \Msun$, and uniform in power-law slope
between $-12$ and $8$.

@@

Theoretical arguments in \citet{Fryer2001} suggest that the black-hole
mass distribution may be well-represented by a decaying exponential
with a minimum mass:
\begin{equation}
  \label{eq:exp-def}
  p(m|\theta) = p(m|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{m}{M_0}
    \right] & \Mmin \leq m \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose again a broad, flat prior
\begin{equation}
  p(\{\Mmin, M_0 \}) = 
  \begin{cases}
    \frac{1}{40^2} & 0 \leq \Mmin, M_0 \leq 40 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
with $\Mmin$ and $M_0$ measured in solar masses.

@@

To look for a second peak in the black-hole mass distribution, we also
examined a two-Gaussian model:
\begin{multline}
  p(m|\theta) = p(m|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right],
\end{multline}
where we restrict $\mu_1 < \mu_2$ with our broad, flat prior:
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    2 \frac{1}{40^4} & 0 \leq \mu_1 \leq \mu_2 \leq 40,
    \quad 0 \leq \sigma_1, \sigma_2 \leq 40, \quad 0 \leq \alpha \leq
    1 \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The one-Gaussian model is a subset of the Gaussian model: when
$\alpha$ is small, then the presence of the first Gaussian is
irrelevant, and the second gives the same mass distribution as a
single Gaussian; when $\alpha$ is large, the reverse situation again
reproduces the single-Gaussian model; and when $\alpha$ is in the
middle of its range, the two means and standard deviations can be
nearly equal, reproducing again the single-Gaussian model.  This is
precisely what we observe from our MCMC samples in the two-Gaussian
model.  In Section \ref{sec:model-selection}, we will quantify more
precisely how much redundancy we have in this model relative to the
one-Gaussian model, but we should emphasize here that there is no
evidence of a significant contribution from a second peak in the
two-Gaussian model.

@@

This work was supported by grant \fixme{XXX}.  Calculations for this
work were performed on the Northwestern Fugu cluster, which was
partially funded by NSF MRI grant \fixme{XXX}.
